Music Classification Using Convolutional Neural Networks

### 2019-12-15 docs/paper.tex
add model and results sections

\usepackage[final]{neurips_2019}
\usepackage[utf8]{inputenc} % allow utf-8 input
\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
\usepackage{hyperref}       % hyperlinks
\usepackage{url}            % simple URL typesetting
\usepackage{booktabs}       % professional-quality tables
\usepackage{amsfonts}       % blackboard math symbols
\usepackage{nicefrac}       % compact symbols for 1/2, etc.
\usepackage{microtype}      % microtypography

@@

\subsection{Style}

Papers to be submitted to NeurIPS 2019 must be prepared according to the
instructions presented here. Papers may only be up to eight pages long,
including figures. Additional pages \emph{containing only acknowledgments and/or
  cited references} are allowed. Papers that exceed eight pages of content
(ignoring references) will not be reviewed, or in any other way considered for
presentation at the conference.

The margins in 2019 are the same as since 2007, which allow for $\sim$$15\%$
more words in the paper compared to earlier years.

Authors are required to use the NeurIPS \LaTeX{} style files obtainable at the
NeurIPS website as indicated below. Please make sure you use the current files
and not previous versions. Tweaking the style files may be grounds for
rejection.

\subsection{Retrieval of style files}

The style files for NeurIPS and other conference information are available on
the World Wide Web at
\begin{center}
  \url{http://www.neurips.cc/}
\end{center}
The file \verb+neurips_2019.pdf+ contains these instructions and illustrates the
various formatting requirements your NeurIPS paper must satisfy.

The only supported style file for NeurIPS 2019 is \verb+neurips_2019.sty+,
rewritten for \LaTeXe{}.  \textbf{Previous style files for \LaTeX{} 2.09,
  Microsoft Word, and RTF are no longer supported!}

The \LaTeX{} style file contains three optional arguments: \verb+final+, which
creates a camera-ready copy, \verb+preprint+, which creates a preprint for
submission to, e.g., arXiv, and \verb+nonatbib+, which will not load the
\verb+natbib+ package for you in case of package clash.

\paragraph{Preprint option}
If you wish to post a preprint of your work online, e.g., on arXiv, using the
NeurIPS style, please use the \verb+preprint+ option. This will create a
nonanonymized version of your work with the text ``Preprint. Work in progress.''
in the footer. This version may be distributed as you see fit. Please \textbf{do
  not} use the \verb+final+ option, which should \textbf{only} be used for
papers accepted to NeurIPS.

At submission time, please omit the \verb+final+ and \verb+preprint+
options. This will anonymize your submission and add line numbers to aid
review. Please do \emph{not} refer to these line numbers in your paper as they
will be removed during generation of camera-ready copies.

The file \verb+neurips_2019.tex+ may be used as a ``shell'' for writing your
paper. All you have to do is replace the author, title, abstract, and text of
the paper with your own.

The formatting instructions contained in these style files are summarized in
Sections \ref{gen_inst}, \ref{headings}, and \ref{others} below.

\section{General formatting instructions}
\label{gen_inst}

The text must be confined within a rectangle 5.5~inches (33~picas) wide and
9~inches (54~picas) long. The left margin is 1.5~inch (9~picas).  Use 10~point
type with a vertical spacing (leading) of 11~points.  Times New Roman is the
preferred typeface throughout, and will be selected for you by default.
Paragraphs are separated by \nicefrac{1}{2}~line space (5.5 points), with no
indentation.

The paper title should be 17~point, initial caps/lower case, bold, centered
between two horizontal rules. The top rule should be 4~points thick and the
bottom rule should be 1~point thick. Allow \nicefrac{1}{4}~inch space above and
below the title to rules. All pages should start at 1~inch (6~picas) from the
top of the page.

For the final version, authors' names are set in boldface, and each name is
centered above the corresponding address. The lead author's name is to be listed
first (left-most), and the co-authors' names (if different address) are set to
follow. If there is only one co-author, list both author and co-author side by
side.

Please pay special attention to the instructions in Section \ref{others}
regarding figures, tables, acknowledgments, and references.

\section{Headings: first level}
\label{headings}

All headings should be lower case (except for first word and proper nouns),
flush left, and bold.

First-level headings should be in 12-point type.

\subsection{Headings: second level}

Second-level headings should be in 10-point type.

\subsubsection{Headings: third level}

Third-level headings should be in 10-point type.

\paragraph{Paragraphs}

There is also a \verb+\paragraph+ command available, which sets the heading in
bold, flush left, and inline with the text, with the heading followed by 1\,em
of space.

\section{Citations, figures, tables, references}
\label{others}

These instructions apply to everyone.

\subsection{Citations within the text}

The \verb+natbib+ package will be loaded for you by default.  Citations may be
author/year or numeric, as long as you maintain internal consistency.  As to the
format of the references themselves, any style is acceptable as long as it is
used consistently.

The documentation for \verb+natbib+ may be found at
\begin{center}
  \url{http://mirrors.ctan.org/macros/latex/contrib/natbib/natnotes.pdf}
\end{center}
Of note is the command \verb+\citet+, which produces citations appropriate for
use in inline text.  For example,
\begin{verbatim}
   \citet{hasselmo} investigated\dots
\end{verbatim}
produces
\begin{quote}
  Hasselmo, et al.\ (1995) investigated\dots
\end{quote}

If you wish to load the \verb+natbib+ package with options, you may add the
following before loading the \verb+neurips_2019+ package:
\begin{verbatim}
   \PassOptionsToPackage{options}{natbib}
\end{verbatim}

If \verb+natbib+ clashes with another package you load, you can add the optional
argument \verb+nonatbib+ when loading the style file:
\begin{verbatim}
   \usepackage[nonatbib]{neurips_2019}
\end{verbatim}

As submission is double blind, refer to your own published work in the third
person. That is, use ``In the previous work of Jones et al.\ [4],'' not ``In our
previous work [4].'' If you cite your other papers that are not widely available
(e.g., a journal paper under review), use anonymous author names in the
citation, e.g., an author of the form ``A.\ Anonymous.''

\subsection{Footnotes}

Footnotes should be used sparingly.  If you do require a footnote, indicate
footnotes with a number\footnote{Sample of the first footnote.} in the
text. Place the footnotes at the bottom of the page on which they appear.
Precede the footnote with a horizontal rule of 2~inches (12~picas).

Note that footnotes are properly typeset \emph{after} punctuation
marks.\footnote{As in this example.}

\subsection{Figures}

\begin{figure}
  \centering
  \fbox{\rule[-.5cm]{0cm}{4cm} \rule[-.5cm]{4cm}{0cm}}
  \caption{Sample figure caption.}
\end{figure}

All artwork must be neat, clean, and legible. Lines should be dark enough for
purposes of reproduction. The figure number and caption always appear after the
figure. Place one line space before the figure caption and one line space after
the figure. The figure caption should be lower case (except for first word and
proper nouns); figures are numbered consecutively.

You may use color figures.  However, it is best for the figure captions and the
paper body to be legible if the paper is printed in either black/white or in
color.

\subsection{Tables}

All tables must be centered, neat, clean and legible.  The table number and
title always appear before the table.  See Table~\ref{sample-table}.

Place one line space before the table title, one line space after the
table title, and one line space after the table. The table title must
be lower case (except for first word and proper nouns); tables are
numbered consecutively.

Note that publication-quality tables \emph{do not contain vertical rules.} We
strongly suggest the use of the \verb+booktabs+ package, which allows for
typesetting high-quality, professional tables:
\begin{center}
  \url{https://www.ctan.org/pkg/booktabs}
\end{center}
This package was used to typeset Table~\ref{sample-table}.

\begin{table}
  \caption{Sample table title}
  \label{sample-table}
  \centering
  \begin{tabular}{lll}
    \toprule
    \multicolumn{2}{c}{Part}                   \\
    \cmidrule(r){1-2}
    Name     & Description     & Size ($\mu$m) \\
    \midrule
    Dendrite & Input terminal  & $\sim$100     \\
    Axon     & Output terminal & $\sim$10      \\
    Soma     & Cell body       & up to $10^6$  \\
    \bottomrule
  \end{tabular}
\end{table}

\section{Final instructions}

Do not change any aspects of the formatting parameters in the style files.  In
particular, do not modify the width or length of the rectangle the text should
fit into, and do not change font sizes (except perhaps in the
\textbf{References} section; see below). Please note that pages should be
numbered.

\section{Preparing PDF files}

Please prepare submission files with paper size ``US Letter,'' and not, for
example, ``A4.''

Fonts were the main cause of problems in the past years. Your PDF file must only
contain Type 1 or Embedded TrueType fonts. Here are a few instructions to
achieve this.

\begin{itemize}

\item You should directly generate PDF files using \verb+pdflatex+.

\item You can check which fonts a PDF files uses.  In Acrobat Reader, select the
  menu Files$>$Document Properties$>$Fonts and select Show All Fonts. You can
  also use the program \verb+pdffonts+ which comes with \verb+xpdf+ and is
  available out-of-the-box on most Linux machines.

\item The IEEE has recommendations for generating PDF files whose fonts are also
  acceptable for NeurIPS. Please see
  \url{http://www.emfield.org/icuwb2010/downloads/IEEE-PDF-SpecV32.pdf}

\item \verb+xfig+ "patterned" shapes are implemented with bitmap fonts.  Use
  "solid" shapes instead.

\item The \verb+\bbold+ package almost always uses bitmap fonts.  You should use
  the equivalent AMS Fonts:
\begin{verbatim}
   \usepackage{amsfonts}
\end{verbatim}
followed by, e.g., \verb+\mathbb{R}+, \verb+\mathbb{N}+, or \verb+\mathbb{C}+
for $\mathbb{R}$, $\mathbb{N}$ or $\mathbb{C}$.  You can also use the following
workaround for reals, natural and complex:
\begin{verbatim}
   \newcommand{\RR}{I\!\!R} %real numbers
   \newcommand{\Nat}{I\!\!N} %natural numbers
   \newcommand{\CC}{I\!\!\!\!C} %complex numbers
\end{verbatim}
Note that \verb+amsfonts+ is automatically loaded by the \verb+amssymb+ package.

\end{itemize}

If your file contains type 3 fonts or non embedded TrueType fonts, we will ask
you to fix it.

\subsection{Margins in \LaTeX{}}

Most of the margin problems come from figures positioned by hand using
\verb+\special+ or other commands. We suggest using the command
\verb+\includegraphics+ from the \verb+graphicx+ package. Always specify the
figure width as a multiple of the line width as in the example below:
\begin{verbatim}
   \usepackage[pdftex]{graphicx} ...
   \includegraphics[width=0.8\linewidth]{myfile.pdf}
\end{verbatim}
See Section 4.4 in the graphics bundle documentation
(\url{http://mirrors.ctan.org/macros/latex/required/graphics/grfguide.pdf})

A number of width problems arise when \LaTeX{} cannot properly hyphenate a
line. Please give LaTeX hyphenation hints using the \verb+\-+ command when
necessary.

\subsubsection*{Acknowledgments}

Use unnumbered third level headings for the acknowledgments. All acknowledgments
go at the end of the paper. Do not include acknowledgments in the anonymized
submission, only in the final paper.

>>>>

\usepackage[final]{neurips_2019}
\usepackage[utf8]{inputenc} % allow utf-8 input
\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
\usepackage{hyperref}       % hyperlinks
\usepackage{url}            % simple URL typesetting
\usepackage{booktabs}       % professional-quality tables
\usepackage{amsfonts}       % blackboard math symbols
\usepackage{nicefrac}       % compact symbols for 1/2, etc.
\usepackage{microtype}      % microtypography
\usepackage{graphicx}
\graphicspath{ {./images/} }

@@

\section{The Model}

\subsection{Trial and Error}

We went through numerous iterations and tried various convolutional neural 
network architectures during development. At first we tried to use convolutional 
windows that would overlap the whole frequency domain and move only in the time domain, 
since although spectrograms can be regarded as arbitrary two dimensional data, the two 
axises have different meanings. This path led nowhere, the model was not able to 
overfit on a small subset of the dataset. We decided to use simple, rectangular convolutional 
windows, which could be called standard by now. This approach was indeed able to overfit, meaning 
it was worth to try learning on the whole dataset. Validation error would cease to decrease 
after the first few epochs. This would be followed by decreasing the input size of the model 
by splitting the samples of 30 seconds length into 4 smaller fragments.

\subsection{The Final Model Architecture}

The final model consists of the following layers:

\begin{itemize}
\item a convolutional layer with a depth of 32 and a window shape of (2, 6)
\item another convolutional layer with a depth of 32 and a window shape of (2, 6)
\item a convolutional layer with a depth of 64 and a window shape identical to the previous ones
\item a fully connected layer of 512 neurons
\end{itemize}

Each convolutional layer was followed by a batch normalization 
layer and a max-pooling layer. L1 and L2 regularizers were adopted to the fully connected layer. 
The model has an output layer of 10 neurons with \emph{softmax} as activation function.

\subsection{Hyperparameters}
\paragraph{Activation functions}
At first we tried using \emph{ReLU} activations between each layer. This lead to many issues, 
since in the context of Mel spectrograms negative numbers have significance. Perceiving this, 
we decided to use \emph{Hyperbolic Tangent} as activation function.

\paragraph{Loss Function}
The loss function used is \emph{Sparse Categorical Crossentropy} since the labels are not one-hot encoded.

\paragraph{Optimizer}
The optimizer algorithm used during training was Stochastic Gradient Descent with a learning rate of 0.01. 
At first we've tried with \emph{Adam} but it did not perform quite as expected. The next choice 
was \emph{Adadelta} which showed promising results, but it did not appear to stable due to the 
fluctuation we have observed on both the graphs of training and validation accuracy.

\paragraph{Number of Epoch and Data Split}
The model was trained for 100 epochs with a batch size of 5. 80 percent of the data was used 
for training, 10 percent was used for validation and the last 10 percent was used for testing.

\paragraph{Vanishing Gradient}

During the training process we've noticed that after a certain number of epochs 
the validation accuracy would cease to converge. This was due to the Vanishing 
Gradient problem, which led to the choice of using \emph{Linear} activation between the 
convolutional layers, leaving \emph{Hyperbolic Tangent} only between the output and 
the fully connected layer.

\section{Results}

\subsection{Training and Validation}

The model is still overfitting with the validation accuracy being stuck at \~60\% and the training accuracy soaring.

\begin{center}
    \includegraphics[scale=0.5]{acc}
		\includegraphics[scale=0.5]{loss}
\end{center}

\subsection{Testing Accuracy}

The final testing accuracy of the model is \textbf{0.61\%}. Below are the Classication Matrix and the final 
Classification Report.

\begin{table}[h]
  \caption{Confusion Matrix}
  \centering
  \begin{tabular}{l | llllllllll}
    Genre & Metal & Classical & Reggae & Jazz & Country & Rock & Hiphop & Disco & Pop & Blues \\
    \midrule
    Metal & 44 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 \\
		Classical & 0 & 28 & 0 & 0 & 1 & 1 & 0 & 0 & 0 & 0 \\
		Reggae & 1 & 1 & 18 & 0 & 1 & 2 & 5 & 3 & 8 & 5 \\
    Jazz & 1 & 11 & 0 & 28 & 1 & 1 & 0 & 0 & 1 & 2 \\
		Country & 0 & 2 & 0 & 3 & 20 & 1 & 0 & 0 & 0 & 4 \\
		Rock & 12 & 0 & 0 & 2 & 5 & 16 & 4 & 2 & 1 & 4 \\
		Hiphop & 7 & 0 & 1 & 0 & 0 & 0 & 28 & 0 & 1 & 2 \\
		Disco & 4 & 1 & 4 & 0 & 1 & 9 & 7 & 8 & 7 & 2 \\
		Pop & 0 & 3 & 1 & 0 & 3 & 2 & 2 & 0 & 35 & 2 \\
		Blues & 1 & 1 & 0 & 5 & 2 & 0 & 0 & 1 & 0 & 20 \\
    \bottomrule
  \end{tabular}
\end{table}

\begin{table}[h]
  \caption{Classification Report}
  \centering
  \begin{tabular}{l | l l l l}
    Genre & Precision & Recall & F1-Score & Support (number of samples) \\
    \midrule
       metal   &    1.00   &   0.63  &    0.77    &    70 \\
   classical    &   0.93    &  0.60  &    0.73    &    47 \\
      reggae    &   0.41    &  0.75   &   0.53    &    24 \\
        jazz    &   0.63    &  0.74   &   0.68    &    39 \\
     country    &   0.67    &  0.59   &   0.62    &    34 \\
        rock    &   0.35    &  0.50   &   0.41    &    32 \\
      hiphop    &   0.72    &  0.61   &   0.66    &    46 \\
       disco    &   0.19    &  0.57   &   0.28    &    14 \\
         pop    &   0.73    &  0.66   &   0.69    &    53 \\
       blues    &   0.67   &   0.49   &   0.56   &     41 \\
    \bottomrule
  \end{tabular}
\end{table}

\begin{table}[h]
  \caption{Classification Report Averages}
  \centering
  \begin{tabular}{l | l l l l}
    Type & Precision & Recall & F1-Score & Support (number of samples) \\
    \midrule
		Macro Average & 0.63 & 0.61 & 0.59 & 400 \\
		Weighted Average & 0.71 & 0.61 & 0.64 & 400 \\
    \bottomrule
  \end{tabular}
\end{table}
